Figures

### 2017-08-22 code/rmarkdown/article.Rmd
fix minor typos and diversity ref

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest, by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes. To achieve this, current methods typically involve partitioning features into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r499; @r500] or phylogenetic trees [@r497]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is conserved in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490]. This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg. @r295].

To overcome this limitation, Faith and Walker (1996) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area. Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; @r204]. Recent work has built on this method using more advanced optimization algorithms [@r482]. However, existing environmental diversity based methods are not often used in conservation planning. This is, in part, because they seek to find the most representative solution that is within a budget, unlike the more commonly used decision support tools (eg. \texttt{Marxan}) that aim to find the cheapest solution subject to a set of targets (or constraints).

@@

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that serve as coordinates in each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.

@@

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

@@

To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies. To assess how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2). We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.

@@

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

@@

By explicitly targeting a representative sample of the species' geographic spread, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig. 3d). This solution prioritized the same number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.

@@

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale-headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using a modified version of the \texttt{spThin R} package; \url{https://github.com/jeffreyhanson/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

@@

# Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF and HPP have Australian Research Council Fellowships. This work was supported by the Centre for Biodiversity and Conservation Science (CBCS). We are also grateful to Dan Rosauer and an anonymous reviewer for their suggestions that greatly improved the manuscript.

>>>>

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest, by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes. To achieve this, current methods typically involve partitioning features into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r499; @r500] or phylogenetic trees [@r497]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is conserved in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups can substantially alter solutions [@r490]. This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg. @r295].

To overcome this limitation, Faith and Walker (1994) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area. Drawing inspiration from the _p_-median problem [@r480], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; @r204]. Recent work has built on this method using more advanced optimization algorithms [@r482]. However, existing environmental diversity based methods are not often used in conservation planning. This is, in part, because they seek to find the most representative solution that is within a budget, unlike the more commonly used decision support tools (eg. \texttt{Marxan}) that aim to find the cheapest solution subject to a set of targets (or constraints).

@@

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that serve as coordinates in each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points in space $s$.

@@

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

@@

To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies. To assess how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2). We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality using \texttt{Gurobi} (version 7.0.2).

@@

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( X, Y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( X, Y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( X, Y \right) \right) &= \text{Max} \begin{cases}
\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

@@

By explicitly targeting a representative sample of the species' geographic spread, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig. 3d). This solution prioritized a similar number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.

@@

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale-headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using a modified version of the \texttt{spThin R} package; \url{www.github.com/jeffreyhanson/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

@@

# Acknowledgements
JOH is supported by an Australian Government Research Training Program (RTP) Scholarship. RAF and HPP have Australian Research Council Fellowships. This work was supported by the Centre for Biodiversity and Conservation Science (CBCS). We are also grateful to Dan Rosauer and an anonymous reviewer for their suggestions that greatly improved the manuscript.

### 2017-08-22 code/rmarkdown/figures.Rmd
figures and readme tweaks
fix minor typos and diversity ref

```{r, echo = FALSE, fig.height = 3.5, fig.width = 2 * column_width, fig.cap = paste0("Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Dark green planning units were selected for protection. Each row of panels show prioritizations generated for a different species. Each column of panels corresponds to a different set of parameters used to generate the prioritization.")}
# prepare data
sim_pus_sel_fpolygons <- plyr::ldply(seq_len(nrow(sim_spp_results)), function(i) {
    curr_data <- sim_pus_fpolygons %>%
                 dplyr::mutate(species = sim_spp_results$species[i],
                               prioritisation =
                                sim_spp_results$prioritisation[i])
    curr_selection <- sim_prioritisations %>%
                      `[[`(as.character(sim_spp_results$species[i])) %>%
                      `[[`(as.character(sim_spp_results$prioritisation[i])) %>%
                      raptr::selections() %>%
                      `[`(as.numeric(curr_data$id))
    curr_data$Selection <-  c("discarded","selected")[curr_selection+1]
    return(curr_data)
})
sim_pus_sel_fpolygons$prioritisation <- factor(
  as.character(sim_pus_sel_fpolygons$prioritisation),
  levels = c("Amount\ntargets", "Amount &\nspace targets",
             "Amount target\nand BLM", "Amount & space\ntargets and BLM"))
# prepare plotting data
sim_letter_data2 <- sim_spp_results %>%
  dplyr::mutate(letter = paste0("(", letters[seq_along(species)],")"),
                long = min(sim_pus_sel_fpolygons$long) + 1.6,
                lat = max(sim_pus_sel_fpolygons$lat) - 1.5)
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_sel_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Selection),
                         color="black") +
  ggplot2::geom_label(data = sim_letter_data2,
                      ggplot2::aes(label = letter, x = long, y = lat),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size=5) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    axis.line.x = ggplot2::element_blank(),
    axis.line.y = ggplot2::element_blank(),
    plot.background = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    panel.background = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    legend.position="none",
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank()) +
  ggplot2::scale_fill_manual(name="Solution",
                             values = c("discarded"="#f7fcfd",
                                        "selected"="#00441b")) +
  ggplot2::facet_grid(species ~ prioritisation)
```

@@

                        fill = "grey85", color = "grey70") +

@@

                        fill="grey85", color="grey70") +

>>>>

```{r, echo = FALSE, fig.height = 3.5, fig.width = 2 * column_width, fig.cap = paste0("Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Dark green planning units were selected for protection. Each row of panels show prioritizations generated for a different species. Each column of panels corresponds to a different set of parameters used to generate the prioritizations.")}
# prepare data
sim_pus_sel_fpolygons <- plyr::ldply(seq_len(nrow(sim_spp_results)), function(i) {
    curr_data <- sim_pus_fpolygons %>%
                 dplyr::mutate(species = sim_spp_results$species[i],
                               prioritisation =
                                sim_spp_results$prioritisation[i])
    curr_selection <- sim_prioritisations %>%
                      `[[`(as.character(sim_spp_results$species[i])) %>%
                      `[[`(as.character(sim_spp_results$prioritisation[i])) %>%
                      raptr::selections() %>%
                      `[`(as.numeric(curr_data$id))
    curr_data$Selection <-  c("discarded","selected")[curr_selection+1]
    return(curr_data)
})
sim_pus_sel_fpolygons$prioritisation <- factor(
  as.character(sim_pus_sel_fpolygons$prioritisation),
  levels = c("Amount\ntargets", "Amount &\nspace targets",
             "Amount target\nand BLM", "Amount & space\ntargets and BLM"))
# prepare plotting data
sim_letter_data2 <- sim_spp_results %>%
  dplyr::mutate(letter = paste0("(", letters[seq_along(species)],")"),
                long = min(sim_pus_sel_fpolygons$long) + 1.6,
                lat = max(sim_pus_sel_fpolygons$lat) - 1.5)
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_sel_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Selection),
                         color="black") +
  ggplot2::geom_label(data = sim_letter_data2,
                      ggplot2::aes(label = letter, x = long, y = lat),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size=5) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    axis.line.x = ggplot2::element_blank(),
    axis.line.y = ggplot2::element_blank(),
    plot.background = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    panel.background = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    legend.position="none",
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank()) +
  ggplot2::scale_fill_manual(name="Solution",
                             values = c("discarded"="#f7fcfd",
                                        "selected"="#00441b")) +
  ggplot2::facet_grid(species ~ prioritisation)
```

@@

                        fill = "grey85", color = "grey70", size = 0.05) +

@@

                        fill="grey85", color="grey70", size = 0.05) +
